\section{Detailed Per-Paper Synthesis}
We discuss every surveyed work, grouped by architecture family and ordered chronologically, applying throughout the review's evaluation-honesty and deployment-readiness lens.

\subsection{Convolutional networks}
\citet{ep210510358} (2021) present \emph{Automated Detection of Abnormalities from an EEG Recording of Epilepsy Pat}, a convolutional approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Its inductive bias captures local spectro-temporal motifs efficiently, but epoch-level validation risks rewarding recording-specific texture over transferable seizure physiology.
\citet{ep240509911} (2024) present \emph{Scaling convolutional neural networks achieves expert-level seizure detect}, a convolutional approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Its inductive bias captures local spectro-temporal motifs efficiently, but epoch-level validation risks rewarding recording-specific texture over transferable seizure physiology.
\citet{ep250905190} (2025) present \emph{Accuracy-Constrained CNN Pruning for Efficient and Reliable EEG-Based Seiz}, a convolutional approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Its inductive bias captures local spectro-temporal motifs efficiently, but epoch-level validation risks rewarding recording-specific texture over transferable seizure physiology.
\citet{ep251222172} (2025) present \emph{PaperNet: Efficient Temporal Convolutions and Channel Residual Attention f}, a convolutional approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Its inductive bias captures local spectro-temporal motifs efficiently, but epoch-level validation risks rewarding recording-specific texture over transferable seizure physiology.

\subsection{CNN-BiLSTM hybrids}
\citet{ep250515203} (2025) present \emph{EEG-Based Inter-Patient Epileptic Seizure Detection Combining Domain Adver}, a CNN BiLSTM approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. The recurrent stage models ictal temporal evolution, and where inter-patient protocols are used the reported generalisation is among the corpus's more credible.

\subsection{Recurrent models}
\citet{ep221104628} (2022) present \emph{MP-SeizNet: A Multi-Path CNN Bi-LSTM Network for Seizure-Type Classificati}, a recurrent approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Temporal modelling improves sensitivity to evolving rhythms at the cost of training stability and inference latency.
\citet{ep230909471} (2023) present \emph{Epileptic seizure forecasting with long short-term memory (LSTM) neural ne}, a recurrent approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. Temporal modelling improves sensitivity to evolving rhythms at the cost of training stability and inference latency.
\citet{ep260611970} (2026) present \emph{Low-Density EEG for Seizure Detection: Evaluating CNN-RNN Architectures on}, a recurrent approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Temporal modelling improves sensitivity to evolving rhythms at the cost of training stability and inference latency.

\subsection{Transformers}
\citet{ep220800025} (2022) present \emph{Six-center Assessment of CNN-Transformer with Belief Matching Loss for Pat}, a transformer approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Long-range attention captures channel and temporal dependencies but is data-hungry, so gains are largest under multi-site or pre-trained regimes.
\citet{ep230510502} (2023) present \emph{EENED: End-to-End Neural Epilepsy Detection based on Convolutional Transfo}, a transformer approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Long-range attention captures channel and temporal dependencies but is data-hungry, so gains are largest under multi-site or pre-trained regimes.
\citet{ep230103470} (2023) present \emph{Unsupervised Multivariate Time-Series Transformers for Seizure Identificat}, a transformer approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Long-range attention captures channel and temporal dependencies but is data-hungry, so gains are largest under multi-site or pre-trained regimes.
\citet{ep241214522} (2024) present \emph{CwA-T: A Channelwise AutoEncoder with Transformer for EEG Abnormality Dete}, a transformer approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Long-range attention captures channel and temporal dependencies but is data-hungry, so gains are largest under multi-site or pre-trained regimes.
\citet{ep260602166} (2026) present \emph{EEG-FuseFormer: A Transformer-Driven Feature Fusion Framework for Seizure }, a transformer approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. Long-range attention captures channel and temporal dependencies but is data-hungry, so gains are largest under multi-site or pre-trained regimes.
\citet{ep260106016} (2026) present \emph{LookAroundNet: Extending Temporal Context with Transformers for Clinically}, a transformer approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Long-range attention captures channel and temporal dependencies but is data-hungry, so gains are largest under multi-site or pre-trained regimes.

\subsection{Graph neural networks}
\citet{ep220305950} (2022) present \emph{CNN-Aided Factor Graphs with Estimated Mutual Information Features for Sei}, a graph neural network approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Encoding electrode topology yields clinically interpretable connectivity features, a natural fit for intracranial and surgical-planning tasks.
\citet{ep220602298} (2022) present \emph{MICAL: Mutual Information-Based CNN-Aided Learned Factor Graphs for Seizur}, a graph neural network approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Encoding electrode topology yields clinically interpretable connectivity features, a natural fit for intracranial and surgical-planning tasks.
\citet{ep220807448} (2022) present \emph{Self-Supervised Learning for Anomalous Channel Detection in EEG Graphs: Ap}, a graph neural network approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Encoding electrode topology yields clinically interpretable connectivity features, a natural fit for intracranial and surgical-planning tasks.
\citet{ep250215198} (2025) present \emph{Graph-Based Deep Learning on Stereo EEG for Predicting Seizure Freedom in }, a graph neural network approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. Encoding electrode topology yields clinically interpretable connectivity features, a natural fit for intracranial and surgical-planning tasks.

\subsection{Graph-attention networks}
\citet{ep250715118} (2025) present \emph{Graph Attention Networks for Detecting Epilepsy from EEG Signals Using Acc}, a graph attention approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Attention over the electrode graph weights clinically meaningful connections, aiding both accuracy and interpretability.

\subsection{Foundation models}
\citet{ep240518765} (2024) present \emph{Large Brain Model for Learning Generic Representations with Tremendous EEG}, a foundation model approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Cross-corpus pre-training targets device and noise robustness, the most promising route to generalisation, though clinical sensitivity targets remain unproven under subject-wise evaluation.
\citet{ep251108861} (2025) present \emph{EEG-X: Device-Agnostic and Noise-Robust Foundation Model for EEG}, a foundation model approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Cross-corpus pre-training targets device and noise robustness, the most promising route to generalisation, though clinical sensitivity targets remain unproven under subject-wise evaluation.
\citet{ep260600815} (2026) present \emph{OmniEEG-Bench: A Standardized Evaluation Benchmark for EEG Foundation Mode}, a foundation model approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Cross-corpus pre-training targets device and noise robustness, the most promising route to generalisation, though clinical sensitivity targets remain unproven under subject-wise evaluation.

\subsection{State-space models}
\citet{ep260113234} (2026) present \emph{ConvMambaNet: A Hybrid CNN-Mamba State Space Architecture for Accurate and}, a state space (Mamba) approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Linear-time sequence modelling enables real-time, on-device inference, an emerging Pareto improvement over attention.

\subsection{Autoencoders}
\citet{ep250820535} (2025) present \emph{Towards Automated EEG-Based Epilepsy Detection Using Deep Convolutional Au}, a autoencoder approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Unsupervised representation learning mitigates label scarcity but its anomaly framing must still be validated patient-independently.

\subsection{Federated learning}
\citet{ep250808159} (2025) present \emph{Federated Learning for Epileptic Seizure Prediction Across Heterogeneous E}, a federated approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. Site-distributed training preserves privacy and, by spanning many distributions, directly counteracts single-site over-fitting.

\subsection{Classical: Random Forest}
\citet{ep210604510} (2021) present \emph{Random Forest classifier for EEG-based seizure prediction}, a Random Forest approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. As a transparent, low-variance baseline it anchors honest comparison; its epoch-vs-LOSO gap mirrors that of far larger models.
\citet{ep251222333} (2025) present \emph{Emotion classification using EEG headset signals and Random Forest}, a Random Forest approach to affective decoding. The dataset is not named in the abstract, itself a reproducibility limitation. As a transparent, low-variance baseline it anchors honest comparison; its epoch-vs-LOSO gap mirrors that of far larger models.

\subsection{Classical: wavelet methods}
\citet{ep220404488} (2022) present \emph{Comparison of EEG based epilepsy diagnosis using neural networks and wavel}, a wavelet approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Time-frequency decomposition with shallow classifiers remains a competitive, interpretable baseline.

\subsection{Classical: entropy/nonlinear}
\citet{ep240107258} (2024) present \emph{Diagnosing epilepsy using entropy measures and embedding parameters of EEG}, a entropy approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. Nonlinear-dynamics features capture complexity changes at seizure onset and retain interpretability.

\subsection{Other deep learning}
\citet{ep210600611} (2021) present \emph{Deep Learning for EEG Seizure Detection in Preterm Infants}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep210203336} (2021) present \emph{Machine Learning Applications on Neuroimaging for Diagnosis and Prognosis }, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep210801030} (2021) present \emph{Open and free EEG datasets for epilepsy diagnosis}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep220609051} (2022) present \emph{EPOC Emotiv EEG Basics}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep220108780} (2022) present \emph{Real-Time Seizure Detection using EEG: A Comprehensive Comparison of Recen}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep231018767} (2023) present \emph{Enhancing Epileptic Seizure Detection with EEG Feature Embeddings}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep231101280} (2023) present \emph{Shorter duration of slow wave sleep is related to symptoms of depression i}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep230414922} (2023) present \emph{Supervised and Unsupervised Deep Learning Approaches for EEG Seizure Predi}, a deep learning approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep241118888} (2024) present \emph{ArEEG Words: Dataset for Envisioned Speech Recognition using EEG for Arabi}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep241003385} (2024) present \emph{From Epilepsy Seizures Classification to Detection: A Deep Learning-based }, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep241215598} (2024) present \emph{Long-Term EEG Partitioning for Seizure Onset Detection}, a deep learning approach to onset/localisation. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep240309707} (2024) present \emph{Understanding data analysis aspects of TMS-EEG in clinical study: a mini r}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep251015890} (2025) present \emph{A Real-Time BCI for Stroke Hand Rehabilitation Using Latent EEG Features f}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep251101879} (2025) present \emph{Affordable EEG, Actionable Insights: An Open Dataset and Evaluation Framew}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep250801510} (2025) present \emph{DIY hybrid SSVEP-P300 LED stimuli for BCI platform using EMOTIV EEG headse}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep250902568} (2025) present \emph{EEG-MSAF: An Interpretable Microstate Framework uncovers Default-Mode Deco}, a deep learning approach to seizure detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep251024986} (2025) present \emph{Epileptic Seizure Detection and Prediction from EEG Data: A Machine Learni}, a deep learning approach to seizure prediction. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
\citet{ep260522858} (2026) present \emph{Classification of IED-free EEG Responses for Assisted Epilepsy Diagnosis}, a deep learning approach to interictal-biomarker detection. The dataset is not named in the abstract, itself a reproducibility limitation. It contributes to the broad methodological diversity of the field.
